---
title: "Oluwaseyi Jimoh Week 3 Assignment"
author: "Oluwaseyi Jimoh"
date: "2024-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Question 1

Using the github link, we import the data from github

```{r}
majors <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
```

Creaty a vector out of the majors column

```{r}
major_vec<- as.vector(majors$Major)
```

From this vector we can run a quick search using | to depict or and ask r to show us majors from the vector maatching DATA or STAtistics


```{r}
grep('DATA|STATISTICS', major_vec, value= TRUE)
```
## Question 2

Adding commas to delineate each string in the list helps r properly create a vector

```{r}
Vegetables <- c("bell pepper", "bilberry",   "blackberry", "blood orange", "blueberry",   "cantaloupe", "chili pepper", "cloudberry",  "elderberry" , "lime",     "lychee",     "mulberry" , "olive" , "salal berry")
```


## Question 3


 (.)\1\1 : This regex will match any expressions as the backslashes are not escaped and the regx is not encased in quotes
 
 "(.)(.)\\2\\1" : This regex will return 4 letters in a palindrome format where the firs 2 letters are repeated backwards e.g abba, otto, iddo
 
 (..)\1 : This regex will match any expressions as the backslashes are not escaped and the regx is not encased in quotes
 
 "(.).\\1.\\1" : This regex will match any expressions where the first letter is repeated 3 times with one character separating each iteration of the letter eg Bababa
 

 "(.)(.)(.).*\\3\\2\\1" : Thie regex will match any expression where the first three letters are repeated at backwards with any number of characters in between them e.g reviver, abccba



## Question 4

```{r}
sample_words <- c("church", "character", "banana","dad", "mom", "eleven", "bababa", "pepper")
```

Regex for words that start and end with the same character

```{r}
str_view(sample_words, "^(.).*\\1$")
```
Regex for words that contain a repeated pair of letters
```{r}
str_view(sample_words, "(..).*\\1")
```
Regex for word that contains one letter in at least 3 places

```{r}
str_view(sample_words, "(a.*){3,}")
```

